Change in Mobility and Site Occupation during the Late Pleistocene in Korea

### 2019-12-03 analysis/paper/paper.Rmd
fixed a few codes to make figures working

```{r cache = TRUE, fig.cap="Map of sites discussed in the text."}
library(here)

library(tidyverse)

@@

mydata <- korean_archaeological_sites

mydata_ages <- 
  mydata %>% 
  separate(`C14(BP)`, into = c('age', 'error'),
          sep = "±") %>% 
  mutate(age_ka = parse_number(age) / 1000,
         error = parse_number(error),
         `altitude(m)_of_main_layer` = parse_number(`altitude(m)_of_main_layer`)) %>% 
  mutate(has_sp = ifelse(!is.na(`SP#`), "yes", "no"))

ggplot(mydata_ages, 
       aes(x = age_ka,
           y = `altitude(m)_of_main_layer`,
       color = as.factor(has_sp))) +
  geom_point(size = 3) +
  xlab("Age of occupation (ka)") +
  ylab("Elevation above sea level (m)") +
  geom_smooth(se = FALSE) +
  scale_colour_discrete(name = "Contains\nstemmed\npoints?") +
  theme_minimal(base_size = 16)

ggsave(here("analysis/figures/site-elevation-by-age.png"))
```

@@

To understand the population change during the appearance of stemmed points, we analyzed the summed probability distributions of 171 radiocarbon dates from 33 cultural layers and it indicates that there is a higher density of radiocarbon dates from around 34 ka. We could cautiously assume that the population flux might relate to the technological innovation.

```{r}
dates_clean <- 
  dates %>% 
  mutate(age = parse_number(as.character(age)),
         error = parse_number(as.character(error))) %>% 
  filter(age < 50000, method=='AMS')

@@

```{r}
#the artifact composition of each sites to figure out what people did as they were occupying the sites.
kasa_long <-
  kasa %>%
  gather(site, 
         count, 
         -X) %>% 
  filter(!is.na(count)) %>% 
  group_by(site) %>% 
  mutate(percentage = count / sum(count, na.rm = TRUE) * 100,
         total = sum(count, na.rm = TRUE)) %>% 
  filter(!is.na(percentage)) %>% 
  # if percentage is <10%, call it 'other'
  mutate(artefact_type = ifelse(as.character(X) == 'stemmed_point',
                                'stemmed_point', 
                                ifelse(percentage >= 10, 
                                as.character(X), "other_tools"))) %>% 
  mutate(axis_label = glue('{site}\n(n = {total})'))  %>% 
  # join to get ages of the sites
  left_join(mydata_ages, 
            by = c('site' = 'site_name')) %>% 
  arrange(age_ka) %>% 
   mutate(age_ka = jitter(age_ka))

>>>>

```{r cache = TRUE, fig.cap="Map of sites discussed in the text."}

@@

mydata_ages <- 
  mydata %>% 
  separate(C14.BP., into = c('age', 'error'),
          sep = "±") %>% 
  mutate(age_ka = parse_number(age) / 1000,
         error = parse_number(error)) %>% 
  mutate(has_sp = ifelse(!is.na(SP.), "yes", "no"))

ggplot(mydata_ages, 
       aes(x = age_ka,
           y = altitude.m._of_main_layer,
       color = as.factor(has_sp))) +
  geom_point(size = 3) +
  xlab("Age of occupation (ka)") +
  ylab("Elevation above sea level (m)") +
  geom_smooth(se = FALSE) +
  scale_colour_discrete(name = "Contains\nstemmed\npoints?") +
  theme_minimal(base_size = 16)

@@

To understand the population change during the appearance of stemmed points, we analyzed the summed probability distributions of 171 radiocarbon dates from 33 cultural layers to get a rough image of population dynamics at the sites. The result indicates that there is a higher density of radiocarbon dates from around 34 ka. We could cautiously assume that the population flux might relate to the technological innovation.

```{r cache=TRUE}
dates_clean <- 
  dates %>% 
  mutate(age = parse_number(as.character(age)),
         error = parse_number(as.character(error))) %>% 
  filter(age < 50000, method=='AMS')

@@

```{r}
#the artifact composition of each sites to figure out what people did as they were occupying the sites.
kasa_long <-
  kasa %>%
  gather(site, 
         count, 
         -X) %>% 
  filter(!is.na(count)) %>% 
  group_by(site) %>% 
  mutate(percentage = count / sum(count, na.rm = TRUE) * 100,
         total = sum(count, na.rm = TRUE)) %>% 
  filter(!is.na(percentage)) %>% 
  # if percentage is <10%, call it 'other'
  mutate(artefact_type = ifelse(as.character(X) == 'stemmed_point',
                                'stemmed_point', 
                                ifelse(percentage >= 10, 
                                as.character(X), "other_tools"))) %>% 
  mutate(axis_label = as.character(glue('{site}\n(n = {total})')))  %>% 
  # join to get ages of the sites
  left_join(mydata_ages, 
            by = c('site' = 'site_name')) %>% 
  arrange(age_ka) #%>% 
  # mutate(age_ka = jitter(age_ka))
